---
title: "Krill Length 48.1 Correlation Analysis with Environmental Drivers"
subtitle: "Alternative Analysis to incorporate in Krill Stock Assessment Model 48.1 SubArea"
author: "Mardones, M; Watters, G.; Cárdenas, C."
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: seaice.bib
csl: apa.csl
link-citations: yes
linkcolor: blue
output:
  html_document:
    keep_md: true
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
editor_options: 
  markdown: 
    wrap: 72
---


```{r setup1}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      dev = 'jpeg',
                      dpi = 300)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```



```{r lib, message=F, echo= TRUE}
library(here)
#analisis
library(ggsignif)
library(ggrepel)
library(ggpubr)
#library(inlmisc)
library(nortest) #para testear distribucion
library(skimr) #provides a frictionless approach to summary statistics 
library(lubridate)
library(easystats) # multiples unciones analiticas
library(lme4)
library(skimr)
library(readxl)
library(fitdistrplus)
# vizualizacion
library(ggridges)
library(sf)
library(GGally)
library(tidyverse, quietly = TRUE)
library(knitr, quietly = TRUE)
library(kableExtra)
library(raster)
library(egg)
library(car) #Variance inflation Factor
library(ggthemes)
library(sjPlot)
library(GGally)
library(CCAMLRGIS)
library(modelsummary)
library(tinytable)
library(psych)
```


```{r}
theme_mau <- function() {
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    panel.background =element_rect(fill = 'aliceblue'),
    panel.border = element_rect(fill = NA, size = 0.6),
    strip.background = element_rect(fill = "white", 
                                    color = "white", 
                                    size = 0.6),
    text = element_text(size = 14),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 90,
                               hjust = 1)
  )
}

```

# Background

The following document intends to carry out a complementary
methodological analysis to correlate environmental variables with the
population dynamics of krill (*Euphausia superba*), in this case, with a
biological component like lengths from fishery monitoring.

# Hypothesis

The primary inquiry pertains to the environmental effects of distinct
physical and oceanographic factors in the Southern Ocean on the krill
population. Our aim is to examine the population structure via krill
length component and determine whether any changes were driven by
environmental factors across any of its dimensions.

# Objective

Once the correlation and effects on the population and/or fishing
indicators on krill have been verified, this analysis aims to have a
time series of the environmental variable to incorporate into the stock
assessment process. Similar work in @Wang2021 but with a longest fishery
history.



# Methodology

## Length structure krill 48 Statistical Subarea

Another important piece of information for a stock evaluation refers to
the biological components such as average sizes and weights across areas
and years. To do this, we will explore the biological data and prepare
the output to add it into stock assessment integrate model
[@Methot2013].

the idea is to correlate mean lengths with SIC.

## Data exploratory analysis

The object `ohbio2` come from data exploration analysis in data request
CCAMLR data. This objetc have bio information from krill.

```{r  echo=FALSE}
#datos entregados por secretaria de CCMLAR
#Data procesada por MMardones
#ohbio <- load("DataLengthKrill.RData")
env_metadata <- new.env()
env_metadata2 <- new.env()
# Cargar cada archivo en su propio entorno
load("~/DOCAS/Data/565_C1_KRI_2021-10-01/DATA_PRODUCT_565.RData", 
     envir = env_metadata)
load("~/DOCAS/Data/669_CHL_2024-03-20/DataProduct/DATA_PRODUCT_669.RData", 
     envir = env_metadata2) # 2020 to 2023
# Extraer los objetos necesarios desde el primer entorno
c1 <- env_metadata$C1
ohbio <- env_metadata$OBS_HAUL_BIOLOGY
# Extraer los objetos necesarios desde el periodo 2020
C1_669 <- env_metadata2$C1
OBS_HAUL_BIOLOGY_669 <- env_metadata2$OBS_HAUL_BIOLOGY
```


Join data set with master as `c1` set. This join is trought
`obs_haul_id` variable to get geoposition variables

```{r warning=FALSE}
ohbio2 <- left_join(c1, ohbio, by="obs_haul_id")%>%
  dplyr::select(-c("notes"))
names(ohbio2)

ohbio2021 <- left_join(C1_669,
                       OBS_HAUL_BIOLOGY_669,
                       by="obs_haul_id") %>%
  dplyr::select(-c("ship_code", "comments" ,"taxon_id"))

names(ohbio2021)

ohbio2merge <- rbind(ohbio2,
                     ohbio2021)

```

Firsts glance. Test how many register have by year. In this case,
`length_total_cm` by season ccamlr. Same exercise in date period
`date_catchperiod_start` to separate dates.

```{r}
ohbio3 <- ohbio2merge %>%
  mutate(Year = year(date_catchperiod_start),
         Month = month(date_catchperiod_start),
         Day = day(date_catchperiod_start))
```


```{r}
names(ohbio3)
ohbio4 <- ohbio3 %>% 
dplyr::select(7, 9, 11, 12, 14, 24, 25, 29, 42, 43, 44, 45, 46, 47)
names(ohbio4)

```

## Plots


Another way to watch this data is regarding thecnics trawl like Figure
\@ref(fig:lenthtrwal).

```{r lenthtrwal, message=F, warning=F, fig.cap="Krill length by Trawl technique"}

trwal <- ggplot(ohbio4 %>% 
               filter(Year>2000),
             aes(x=length_total_cm, 
                 y = as.factor(Year),
                 fill= trawl_technique))+
  #geom_joy(alpha=0.9) +
  geom_density_ridges(stat = "binline", bins = 50, 
                      scale = 1.8, 
                      draw_baseline = FALSE,
                      alpha=0.5)+
  facet_wrap(.~asd_code, ncol=3) +   
  geom_vline(xintercept = 3.6, color = "red")+
  scale_x_continuous(breaks = seq(from = 1, to = 10, 
                                  by = 1))+
  scale_y_discrete(breaks = seq(from = 2000, 
                                to = 2023, by = 1))+
  scale_fill_viridis_d(name="Trawl Technique",
                       option="F")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))+
  xlim(0,10)+
  xlab("Longitud (cm.)")+
  ylab("")
trwal
```





## Mean Length

Recruit estimate (% Low Recruit length) [@Perry2020]

Un diaggrama de caja para el porcentage de individuos bajo talla (3.6
mmm.)

```{r warning=FALSE}
box <- ggplot(ohbio4) +
  geom_boxplot(aes(length_total_cm, 
                   group=season_ccamlr), 
               alpha=0.3,
               fill=3)+
  geom_vline(xintercept = 3.6, color = "red")+
  xlim(0, 7)+
  coord_flip()+
  facet_wrap(~asd_code, ncol=1)+
  theme_bw()+
  labs(x="% Below Recruit length (3.6 mm.)",
       x="")
hi <- ggplot(ohbio4)+
  geom_histogram(aes(y=length_total_cm), 
                 fill=3,
                 alpha=0.3,
                 color="black")+
  coord_flip()+
  theme_bw()+
  ylim(0,10)

ggarrange(box, hi, ncol = 2)
  
```
## Preparo  bases para toda la 48 con `ohbio4`

```{r}
# 48.1
s481 <- ohbio4 %>% 
  drop_na(length_total_cm) %>% 
  filter(asd_code== "481",
         vessel_nationality_code != "CHN")  
# cut in order
s481$catlon <- cut(x = s481$length_total_cm, 
                 breaks = seq(0,7,0.2),
                 labels = seq(0,6.8,0.2),
                 right = FALSE)

s481t <- table(s481$Year, s481$catlon)
s481t1 <-t(as.data.frame.matrix(s481t)); class(s481t1)


# 48.2
s482 <- ohbio4 %>% 
  drop_na(length_total_cm) %>% 
  filter(asd_code== "482",
         vessel_nationality_code != "CHN")  
# cut in order
s482$catlon <- cut(x = s482$length_total_cm, 
                 breaks = seq(0,7,0.2),
                 labels = seq(0,6.8,0.2),
                 right = FALSE)

s482t <- table(s482$Year, s482$catlon)
s482t1 <-t(as.data.frame.matrix(s482t)); class(s482t1)

# 48.3
s483 <- ohbio4 %>% 
  drop_na(length_total_cm) %>% 
  filter(asd_code== "483",
         vessel_nationality_code != "CHN")  
# cut in order
s483$catlon <- cut(x = s483$length_total_cm, 
                 breaks = seq(0,7,0.2),
                 labels = seq(0,6.8,0.2),
                 right = FALSE)

s483t <- table(s483$Year, s483$catlon)
s483t1 <-t(as.data.frame.matrix(s483t)); class(s483t1)

# A su vez puedo generar el archivo por separado
write.csv(s481t1, "lenght481_nochina.csv", sep = ",", row.names = TRUE)
write.csv(s482t1, "lenght482_nochina.csv", sep = ",", row.names = TRUE)
write.csv(s483t1, "lenght483_nochina.csv", sep = ",", row.names = TRUE)


```

(Probar datos con RaadTools library [@Raadtools2022])

## Maps length data

First thing is get different rater layer to join krill data length
according different porpoises.

```{r raster}
# Cargo linea de costa
coast <- load_Coastline()
coast1<- st_as_sf(coast) 
coast2 = st_transform(coast1, "+proj=latlong +ellps=WGS84")
# con SSMU
ssmu <- load_SSMUs()
ssmu481 <- subset(ssmu[c(2,3,4,5,6,7,17),])
ssmu481a <- st_as_sf(ssmu481) 
ssmu481aa = st_transform(ssmu481a, "+proj=latlong +ellps=WGS84")

# con Statistical Areas con foco en 48.1
suba <- load_ASDs()
suba1 <- subset(suba[(3),])
suba1a<- st_as_sf(suba1) 
suba1aa = st_transform(suba1a, "+proj=latlong +ellps=WGS84")


# Uso las agrupaciones de DMP1
dmp1 <- st_read("~/DOCAS/Mapas/Antarctic_SHPfiles/D1MPA-model-2019.shp",
                quiet=T)
dmp1=st_transform(dmp1, 6932)

# Uso las agrupaciones de Strata
strata <- st_read("~/DOCAS/Mapas/Antarctic_SHPfiles/Strata.shp",
                quiet=T)
strata=st_transform(strata,
                    "+proj=latlong +ellps=WGS84")

# Uso las agrupaciones de Strata Lucas HW
strataclip <- st_read("~/DOCAS/Data/WAP-Heatwaves/peninsula_correct.shp",
                quiet=T)
strataclip=st_transform(strata, 
                        "+proj=latlong +ellps=WGS84")


```


```{r maptest, fig.cap="Strata Maps in 48.1"}
# y testeo el mapa
ssmap <- ggplot()+
  geom_sf(data = strata %>%
           filter(ID != "Outer")%>% 
                 mutate(ID = str_replace_all(ID, "Extra", "GS")), 
          aes(fill=ID, 
              alpha=0.3))+
  geom_text_repel(data = strata %>% 
            filter(ID != "Outer"), 
            aes(x = Labx, y = Laby, 
                label = ID), 
            min.segment.length = 2,
                             box.padding = 2,
                             max.overlaps = 15)+
  geom_sf(data = coast2, colour="black", fill=NA)+
  geom_sf(data= suba1aa, 
          colour="red",
          fill=NA)+
  scale_fill_viridis_d(option = "F")+
  ylim(230000, 2220000)+
  xlim(-3095349 , -1858911)+
  # coord_sf(crs = 32610)+ #sistema de prpyecccion para campos completos
  coord_sf(crs = 6932)+
  theme_bw()+
  theme(legend.position = "none")
ssmap
```

## Grouping Length data into Grid

### SSMU 48.1 CCAMLR

### SACO CHINA!!! y uso solo hasta 2020 para el analisis

```{r ssmu1}
dim(ohbio4)
ohbio5 <- ohbio4 %>% 
  filter(asd_code==481,
         Year < 2021,
         vessel_nationality_code != "CHN") %>% 
  dplyr::select(1, 6, 7, 8, 9, 10, 11, 12) 

ohbio5_conchina<- ohbio4 %>% 
  filter(asd_code==481,
         Year < 2021,) %>% 
  dplyr::select(1, 6, 7, 8, 9, 10, 11, 12) 

ohbio6 <- st_as_sf(ohbio5 %>% 
                     drop_na(latitude_set_end), 
                   coords = c("longitude_set_end", 
                              "latitude_set_end"),  
                  crs = "+proj=latlong +ellps=WGS84")
ohbio6_conchina <- st_as_sf(ohbio5_conchina %>% 
                     drop_na(latitude_set_end), 
                   coords = c("longitude_set_end", 
                              "latitude_set_end"),  
                  crs = "+proj=latlong +ellps=WGS84")
```

This grid has the same characteristics as the environmental data grids
that will be called up later. This grid is 1x0.5 degrees which allows a
clear visualization of the processes, whether biological and/or
environmental.

```{r gri}

Grid<- suba1aa  %>% #pm481 es el plot base original linea 481
  sf::st_make_grid(cellsize = c(1,0.5)) %>% # para que quede cuadrada
  sf::st_cast("MULTIPOLYGON") %>%
  sf::st_sf()  %>%  # objeto en spatial feature
  dplyr::mutate(cellid = row_number()) 

# Clean the input data by removing duplicate vertices and making the object topologically valid
grid3 <- st_make_valid(Grid)

# Corto la grilla dentro de las SSMU
gridcrop1 <- tmaptools::crop_shape(grid3, 
                                   suba1aa,
                                   polygon = TRUE)

# the first object drives the output geometry
grilen <- grid3 %>%
  st_join(ohbio6) %>% 
  group_by(cellid)

names(grilen)
```

calculando algunos indicadores

```{r plogri}
grilen3 <- grilen %>%
  drop_na(length_total_cm) %>% 
  dplyr::group_by(cellid,Year) %>% 
  dplyr:: summarise(
    lepro = mean(length_total_cm, na.rm = TRUE),  # Media
    sd_lepro = sd(length_total_cm, na.rm = TRUE) # Desviación estándar
  )
```


solo la grilla

```{r}
glvoid <- ggplot() +
  geom_sf(data = coast2, colour="black", fill=NA)+
  geom_sf(data = gridcrop1, colour="black", fill=NA)+
  geom_sf(data= suba1aa, fill=NA, col="red")+
  theme(panel.background = element_rect(fill = 'aliceblue'),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
   ylim(230000, 2220000)+
  xlim(-3095349 , -1858911)+
  # coord_sf(crs = 32610)+ #sistema de prpyecccion para campos completos
  coord_sf(crs = 6932)+
  scale_alpha(guide="none")+
  theme_few()
glvoid
```


Pro
```{r plogri2}
glpro <- ggplot() +
   geom_sf(data=grilen3 %>% 
             drop_na(lepro) %>% 
             mutate(lepro_nivel =cut(na.omit(lepro),
                   breaks = seq(2, 6, by = 0.5),
                        labels = c("2-2.5", "2.5-3.0", "3.0-3.5", "3.5-4.0",
                                   "4.0-4.5", "4.5-5.0", "5.0-5.5", "5.5-6.0"),
                        include.lowest = TRUE)),
           aes(fill = lepro_nivel), color=NA) +
  scale_fill_brewer(palette = "YlOrBr" ,
                    name = "")+
  geom_sf(data = suba1aa, fill=NA, col=2)+
  geom_sf(data = coast2, 
          colour=NA, 
          fill="grey",
          alpha=0.5)+
  #facet_wrap(~Year, ncol=6)+
  theme_mau()+
  # ylim(-70, -60)+
  # xlim(-70, -50)
  # another kind projection
  #coord_sf(crs = 32610)+ #sistema de prpyecccion para campos completos
  ylim(230000, 2220000)+
  xlim(-3095349 , -1858911)+
  labs(subtitle = "Krill Length mean (cm)")+
  coord_sf(crs = 6932)
gl <- ggplot() +
   geom_sf(data=grilen3 %>% 
             drop_na(sd_lepro) %>% 
             mutate(sd_nivel =cut(na.omit(sd_lepro),
                   breaks = seq(0, 1.2, by = 0.2), 
                   labels = c("0-0.2", "0.2-0.4", "0.4-0.6", "0.6-0.8", "0.8-1.0", "1.0-1.2"),
                   include.lowest = TRUE)),
           aes(fill = sd_nivel), color=NA) +
  scale_fill_brewer(palette = "YlOrBr" ,
                    name = "")+
  geom_sf(data = suba1aa, fill=NA, col=2)+
  geom_sf(data = coast2, 
          colour=NA, 
          fill="grey",
          alpha=0.5)+
  #facet_wrap(~Year, ncol=6)+
  theme_mau()+
  #theme(legend.position = "bottom")+
  # ylim(-70, -60)+
  # xlim(-70, -50)
  # another kind projection
  #coord_sf(crs = 32610)+ #sistema de prpyecccion para campos completos
  ylim(230000, 2220000)+
  xlim(-3095349 , -1858911)+
  labs(subtitle = "Krill Length Standar Deviation")+
  coord_sf(crs = 6932)

ggarrange(glpro, gl,
          ncol=2) 
```


###  STRATA

Length composition by Strata CCAMLR to visualization first. First step is group data into to poligons strata.

```{r wranlingdata}
class(ohbio6)
dim(ohbio6)
glimpse(ohbio6)

class(strata)
glimpse(strata)
dim(strata)


ohbiopro <- ohbio6 %>%
  dplyr::group_by(gear_type, Year, Month, geometry) %>% 
  dplyr::summarize(tapro = mean(length_total_cm))

# comoprobar si tengo datos duplicados
strata2 <- st_make_valid(strata)
ohbio62 <- st_make_valid(ohbio6)
ohbio6_conchina2 <- st_make_valid(ohbio6_conchina)


sf3 <- st_join(strata2, ohbiopro)
sf4 <- st_join(strata2, ohbio6)
sf5 <- st_join(strata2,
               ohbio6_conchina2)
dim(sf4)
dim(sf3)
dim(sf5)
```


```{r saveda, eval=FALSE}
#save("sf4", file = "sf4_nochina.RData")
#save("sf3", file = "sf3_mochina.RData")
#save("sf5", file = "sf5_china.RData")
load("~/DOCAS/Data/Krill_Length_Cor/sf4_nochina.RData")
```



histogram length data


```{r}
jzstrata <- ggplot(sf4 %>% 
               filter(Year>2000,
                      ID !="Outer") %>% 
                 mutate(ID = str_replace(ID, "Extra", "GS")),
             aes(x=length_total_cm, 
                 y = as.factor(Year), 
                 fill=ID))+
  geom_density_ridges(stat = "binline", bins = 30, 
                      scale = 1.9, 
                      draw_baseline = FALSE,
                      alpha=0.9)+
  facet_wrap(.~ID, ncol=7) +   
  geom_vline(xintercept = 3.6, color = "red")+
  scale_x_continuous(breaks = seq(from = 1, to = 10, 
                                  by = 2))+
  scale_y_discrete(breaks = seq(from = 2000, 
                                to = 2020, by = 1))+
  scale_fill_viridis_d(name="Strata",
                       option="F")+
  theme_minimal()+
  theme(legend.position="none")+
  xlab("Length (cm)")+
  ylab("")
jzstrata
```

Violin plot by strata with China

```{r}

viostrachina <- ggplot(sf5 %>% 
               filter(Year>2000,
                      ID !="Outer")%>% 
                 mutate(ID = str_replace(ID, "Extra", "GS")),
             aes(x=length_total_cm, 
                 y = as.factor(Year), 
                 fill=ID))+
  scale_fill_viridis_d(name="Strata",
                       option="F")+
  geom_violin(alpha=0.8,
              color="NA")+
  geom_boxplot(width=0.1,
               outlier.shape = NA,
               show.legend = FALSE,
               alpha=0.1)+
  geom_vline(xintercept = 3.6, color = "red")+
  facet_wrap(.~ID, ncol=7) + 
  theme_minimal()+
  theme(legend.position="none")+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))+
  ylab("")+
  xlab("Length (cm)")+
  coord_flip()

viostra <- ggplot(sf5 %>% 
               filter(Year>2000,
                      ID !="Outer",
                      vessel_nationality_code != "CHN")%>% 
                 mutate(ID = str_replace(ID, "Extra", "GS")),
             aes(x=length_total_cm, 
                 y = as.factor(Year), 
                 fill=ID))+
  scale_fill_viridis_d(name="Strata",
                       option="F")+
  geom_violin(alpha=0.8,
              color="NA")+
  geom_boxplot(width=0.1,
               outlier.shape = NA,
               show.legend = FALSE,
               alpha=0.1)+
  geom_vline(xintercept = 3.6, color = "red")+
  facet_wrap(.~ID, ncol=7) + 
  theme_minimal()+
  theme(legend.position="none")+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))+
  ylab("")+
  xlab("Length (cm)")+
  coord_flip()



ggarrange(viostrachina,
          viostra,
          ncol = 1)

```

```{r}
# Calculando datos con China
sf5_means_with_china <- sf5 %>%
  filter(Year > 2000, ID != "Outer") %>%
  mutate(ID = str_replace(ID, "Extra", "GS"),
         Data_Source = "With China") %>% # Etiqueta para identificar el origen
  group_by(Year, ID, Data_Source) %>%
  summarize(mean_length = mean(length_total_cm, na.rm = TRUE),
            sd_length = sd(length_total_cm, na.rm = TRUE)) %>%
  ungroup()

# Calculando datos sin China
sf5_means_no_china <- sf5 %>%
  filter(Year > 2000, ID != "Outer", vessel_nationality_code != "CHN") %>%
  mutate(ID = str_replace(ID, "Extra", "GS"),
         Data_Source = "Without China") %>% # Etiqueta para identificar el origen
  group_by(Year, ID, Data_Source) %>%
  summarize(mean_length = mean(length_total_cm, na.rm = TRUE),
            sd_length = sd(length_total_cm, na.rm = TRUE)) %>%
  ungroup()

# Combinando ambos conjuntos de datos
sf5_combined <- bind_rows(sf5_means_with_china, sf5_means_no_china)

# Gráfico combinado
trend_combined <- ggplot(sf5_combined, aes(x = Year, 
                                           y = mean_length, 
                                           color = ID, 
                                           linetype = Data_Source)) +
  geom_line(size = 1.2) +
  geom_point(aes(shape = Data_Source), size = 3) +
  geom_errorbar(aes(ymin = mean_length - sd_length, 
                    ymax = mean_length + sd_length),
                width = 0.2, alpha = 0.5) +
  scale_color_viridis_d(name = "Strata", option = "F") +
  scale_linetype_manual(name = "Data Source", values = c("With China" = "solid", "Without China" = "dashed")) +
  scale_shape_manual(name = "Data Source", values = c("With China" = 16, "Without China" = 17)) +
  labs(x = "Year", y = "Mean Length (cm)",
       title = "Trends in Mean Length by Strata (With and Without China)") +
  theme_few() +
  theme(legend.position = "bottom",
        legend.box = "horizontal")

# Mostrar el gráfico
trend_combined

# Mostrar los gráficos

```
test

```{r}
# Filtrando datos para asegurar el mismo rango de años
paired_data <- sf5_combined %>% 
  filter(Data_Source %in% c("With China", "Without China")) %>%
  pivot_wider(names_from = Data_Source, values_from = mean_length) %>%
  filter(!is.na(`With China`) & !is.na(`Without China`))
```


```{r}

# Ajuste del modelo lineal mixto
lmm <- lmer(mean_length ~ Data_Source + (1 | Year) + (1 | ID), data = sf5_combined)

# Resumen del modelo
summary(lmm)
anova(lmm)

# T-Test pareado
# T-Test independiente
independent_t_test <- t.test(mean_length ~ Data_Source, data = sf5_combined)
print(independent_t_test)
```

color of nation without china

```{r}
color <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a','#ffff99')
colnat <- ggplot() +
  geom_sf(data = strata, fill=NA, col=2)+
  geom_sf(data = coast2, colour="#636363", 
          fill="white")+
   geom_sf(data=ohbio6_conchina %>% 
               filter(Year>2000,
                      vessel_nationality_code == "CHN"),
           aes(color = as.factor(Year))) +
  #facet_wrap(~Year, ncol=10)+
  scale_color_manual(values = color)+
  theme(panel.background = element_rect(fill = 'aliceblue'),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
  # ylim(-70, -60)+
  # xlim(-70, -50)
  # another kind projection
  ylim(230000, 2220000)+
  xlim(-3095349 , -1858911)+
  #coord_sf(crs = 32610)+ #sistema de prpyecccion para campos completos
  coord_sf(crs = 6932)+
  labs(title = "Operation by year by china Fleet")+
  theme_few()
colnat
```
maps length

```{r}
glstrata <- ggplot() +
   geom_sf(data=sf3%>% 
             drop_na(tapro) %>% 
               filter(Year>1999,
                      ID !="Outer") %>% 
             mutate(tapro_nivel =cut(na.omit(tapro),
                   breaks = seq(2, 6, by = 1),
                        labels = c("2-3", "3-4", "4-5", "5-6"),
                        include.lowest = TRUE)),
           aes(fill = tapro_nivel), 
           color=NA) +
   scale_fill_brewer(palette = "YlOrBr" ,
                    name = "(cm)")+
  geom_sf(data = strata, fill=NA, col=2)+
  geom_sf(data = coast2, colour="#636363", 
          fill="white")+
  facet_wrap(~Year, ncol=6)+
  theme(panel.background = element_rect(fill = 'aliceblue'),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
  # ylim(-70, -60)+
  # xlim(-70, -50)
  # another kind projection
  ylim(230000, 2220000)+
  xlim(-3095349 , -1858911)+
  #coord_sf(crs = 32610)+ #sistema de prpyecccion para campos completos
  coord_sf(crs = 6932)+
  labs(title = "Length Krill by Strata and Year")+
  theme_mau()
glstrata
```

Prepare template length data by strata for posterior analysis like "Intrinsic Productivity". We can found this in this [link](https://github.com/MauroMardones/LBSPR_Krill)

#### Grouping by strata

```{r eval=FALSE, echo=TRUE}
# Maybe this could be a loop , but i didnt find the way...
# Elephan Island
ei <- sf4 %>% 
  drop_na(length_total_cm) %>% 
  filter(ID== "EI")  
# cut in order
ei$catlon <- cut(x = ei$length_total_cm, 
                 breaks = seq(0,7,0.2),
                 labels = seq(0,6.8,0.2),
                 right = FALSE)

eit <- table(ei$Year, ei$catlon)
eit1 <-t(as.data.frame.matrix(eit)); class(eit1)


# SSIW
ssiw <- sf4 %>% 
  drop_na(length_total_cm) %>% 
  filter(ID== "SSIW")  
# cut in order
ssiw$catlon <- cut(x = ssiw$length_total_cm, 
                 breaks = seq(0,7,0.2),
                 labels = seq(0,6.8,0.2),
                 right = FALSE)

ssiwt <- table(ssiw$Year, ssiw$catlon)
ssiwt1 <-t(as.data.frame.matrix(ssiwt)); class(ssiwt1)

# BS
bs <- sf4 %>% 
  drop_na(length_total_cm) %>% 
  filter(ID== "BS")  
# cut in order
bs$catlon <- cut(x = bs$length_total_cm, 
                 breaks = seq(0,7,0.2),
                 labels = seq(0,6.8,0.2),
                 right = FALSE)

bst <- table(bs$Year, bs$catlon)
bst1 <-t(as.data.frame.matrix(bst)); class(bst1)

# JOIN  (very poor data, just a couple years)
join <- sf4 %>% 
  drop_na(length_total_cm) %>% 
  filter(ID== "JOIN")  
# cut in order
join$catlon <- cut(x = join$length_total_cm, 
                 breaks = seq(0,7,0.2),
                 labels = seq(0,6.8,0.2),
                 right = FALSE)

joint <- table(join$Year, join$catlon)
joint1 <-t(as.data.frame.matrix(joint)); class(joint1)

# Extra
extra <- sf4 %>% 
  drop_na(length_total_cm) %>% 
  filter(ID== "Extra")  
# cut in order
extra$catlon <- cut(x = extra$length_total_cm, 
                 breaks = seq(0,7,0.2),
                 labels = seq(0,6.8,0.2),
                 right = FALSE)

extrat <- table(extra$Year, extra$catlon)
extrat1 <-t(as.data.frame.matrix(extrat)); class(extrat1)
```

Generate the files to use in intrinsic productivity analysis.

```{r eval=FALSE, echo=TRUE, warning=FALSE}
# A su vez puedo generar el archivo por separado
write.csv(eit1, "lenghtEI_nochina.csv", sep = ",", row.names = TRUE)
write.csv(ssiwt1, "lenghtSSIW_nochina.csv", sep = ",", row.names = TRUE)
write.csv(bst1, "lenghtBS_nochina.csv", sep = ",", row.names = TRUE)
write.csv(joint1, "lenghtJOIN_nochina.csv", sep = ",", row.names = TRUE)
write.csv(extrat1, "lenghtExtra_nochina.csv", sep = ",", row.names = TRUE)

```


## Environmental data

Load environmental data to merge with length.


```{r echo=FALSE}
load("~/DOCAS/Data/KrillLows/DataEnvKrill.RData")
```

Get data specific object

```{r include=FALSE}
ls()
get("dataenvf")
```

Join length and environental data by `cellid` and `Year` with a reduce DB.

```{r}
class(dataenvf)

envgru <- dataenvf %>%
  dplyr::select(1, 2, 3, 6, 9) %>% 
  drop_na(ANO.x) %>% 
  dplyr::rename(Year=ANO.x) %>% 
  dplyr::group_by(cellid,Year) %>% 
  dplyr::summarise(meansic2=mean(meansic),
                   meanchl2=mean(meanchl),
                   meantsm2=mean(meantsm))
dim(envgru)
summary(envgru)
# 
# # Test
# envgru2 <- dataenvf %>%
#   dplyr::select(1, 2, 3, 6, 9, 10) %>%
#   drop_na(ANO.x) %>%
#   dplyr::rename(Year=ANO.x) %>%
#   dplyr::group_by(cellid,Year, geometry) %>%
#   dplyr::summarise(meansic2=mean(meansic),
#                    meanchl2=mean(meanchl),
#                    meantsm2=mean(meantsm))
```

Join merge object in a new data base.

```{r}
names(grilen3)
names(envgru)
envlen <- merge(grilen3, envgru, 
                 by = c("cellid", "Year"))
# change names

envname <- envlen %>% 
  dplyr::rename(Length=lepro,
         SIC=meansic2,
         CHla=meanchl2,
         SSM=meantsm2)
head(envname)
```

# Results

## GLM Length-Environmental

Debo trabajar ciertos componentes de los datos para poder hacer el
analisis bajo supuestos estructurales.

-   Comprobar distribución de Longitudes medias de krill
-   Cambiar a factor `Year` y `cellid`
-   Filtrar datos `NA`
-   Filtrar outliers en datos ambientales

### Test Correlation

Test de comportamiento de la variable de longitudes `lenpro` mediante 2
test; 1) Shapiro-Wilk `shapiro.test()` y Kolmogorov-Smirnov `ks.test()`

```{r}
shapiro.test(envname$Length)
ks.test(envname$Length, "pnorm", mean(envname$Length), sd(envname$Length))
tal<-ggplot(envname, aes(x=Length)) + 
  geom_histogram(color="red", fill="white")+
  theme_minimal()+
  labs(x="Length Mean Krill (cm.)",
       y="Frecuency")
tal
```

### Cambiar a factor `Year` y `cellid`

```{r}
envname$Year<-as.factor(envname$Year)
envname$cellid<-as.factor(envname$cellid)
glimpse(envname)
```

### Filtrar datos `NA` if we have. Test

```{r}
# Are there missing values?
colSums(is.na(envname))
```

### Filter outliers environmental and biological data

Test trought plot all variables in `envlen` object

```{r}
boxsic <- ggplot(envname)+
  geom_boxplot(aes(Year, SIC),
               outlier.colour = "red")+
  ylim(0,50)+
  theme_minimal()
boxtsm <- ggplot(envname)+
  geom_boxplot(aes(Year, SSM-270),
               outlier.colour = "red")+
  labs(y="SST")+
  theme_minimal()
boxchl <- ggplot(envname)+
  geom_boxplot(aes(Year, CHla),
               outlier.colour = "red")+
  theme_minimal()
ggarrange(boxsic, boxtsm, boxchl, nrow = 3)

```

is necessary get off outliers from Sea ice index. Get environmental
outliers data and again...

```{r warning=F, message=FALSE}
envlen2 <- envlen %>% 
  filter(meansic2<100)

#plot 2
         
boxsic2 <- ggplot(envlen2)+
  geom_boxplot(aes(Year, meansic2),
               outlier.colour = "red")+
  theme_minimal()
boxtsm2 <- ggplot(envlen2)+
  geom_boxplot(aes(Year, meantsm2),
               outlier.colour = "red")+
  theme_minimal()
boxchl2 <- ggplot(envlen2)+
  geom_boxplot(aes(Year, meanchl2),
               outlier.colour = "red")+
  theme_minimal()

p<-ggplot(envlen2, aes(x=meansic2)) + 
  geom_histogram(color="black", fill="white")+
  theme_minimal()

q<-ggplot(envlen2, aes(x=meantsm2)) + 
  geom_histogram(color="black", fill="white")+
  theme_minimal()

t<-ggplot(envlen2, aes(x=meanchl2)) + 
  geom_histogram(color="black", fill="white")+
  theme_minimal()

ggarrange(boxsic2, boxtsm2, boxchl2, p, q, t, nrow = 2, ncol=3)
```

## Correlation Analysis

## Simple correlation identification toget information about variables

Firts filter in our dataframe `envlen3`.

Regarding difference between `Spearman` and `pearson`, both, the
Spearman test and the Pearson test are statistical methods used to
assess the correlation between two variables. The main difference
between them is that the Pearson test evaluates the linear correlation
between two continuous variables, while the Spearman test evaluates the
monotonic correlation between two continuous or order variables.

In the case of the Pearson test, the degree of association between two
continuous variables is measured through a correlation coefficient that
varies between -1 and 1. A value of 1 indicates a perfectly positive
correlation, a value of -1 indicates a perfectly negative correlation,
and a value of 0 indicates no correlation between the two variables.

On the other hand, the Spearman test is based on the range of the
variables, instead of the actual values. In other words, this test
evaluates the correlation between two order variables, where the values
of each variable are ranked from lowest to highest, and ranges are used
instead of actual values. Spearman's correlation coefficient also varies
between -1 and 1, but it measures the monotonic correlation between two
variables, that is, if one variable increases, the other variable also
increases or decreases[@McCulloch2001].

In summary, the main difference between the Spearman test and the
Pearson test is that the former is used to assess the monotonic
correlation between two order variables, while the latter is used to
assess the linear correlation between two continuous variables.

-   Spearman's Rank Correlation Coefficient

This coefficient is used to see if there is any significant relationship
between the two datasets, and operates under the assumption that the
data being used is ordinal, which here means that the numbers do not
indicate quantity, but rather they signify a position of place of the
subject's standing (e.g. 1st, 2nd, 3rd, etc.)

$\begin{aligned} r_s = \frac{\sum_{i=1}^n (x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum_{i=1}^n (x_i - \bar{x})^2}\sqrt{\sum_{i=1}^n (y_i - \bar{y})^2}} \end{aligned}$

-   Pearson Product-Moment Coefficient

This is the most widely used correlation analysis formula, which
measures the strength of the 'linear' relationships between the raw data
from both variables, rather than their ranks. This is an dimensionless
coefficient, meaning that there are no data-related boundaries to be
considered when conducting analyses with this formula, which is a reason
why this coefficient is the first formula researchers try.

$\begin{aligned} r = 1- \frac{6\sum_{i=1}^n D_{i}^n}{n (n^2 - 1)}\end{aligned}$

(Is posible report outpus by @Report2023)

```{r}
# change to data frame
envlen3 <- as.data.frame(envlen2)
envlen4 <- envlen3 %>% 
  dplyr::select(3, 4, 5, 6, 7)
```

Plot to show correlation trough different ways. First, the matri
correlation

```{r}
kableExtra::kable(cor(envlen4)) %>% 
  kable_styling()
# covariance matrix
kableExtra::kable(cov(envlen4)) %>% 
  kable_styling()
```


```{r message=FALSE, warning=FALSE}
pairs.panels(envlen4 %>% 
               mutate(meantsm2=meantsm2-270)%>% 
               dplyr::rename(Length=lepro,
                             SDLength = sd_lepro,
         SIC=meansic2,
         Chla=meanchl2,
         SST=meantsm2),
             smooth = TRUE,      
             scale = FALSE,      
             density = TRUE,     
             ellipses = TRUE,  
             method = "pearson",
             pch = 21,           
             lm = FALSE,        
             cor = TRUE,         
             jiggle = FALSE,     
             factor = 2,        
             hist.col = 4,       # Color de los histogramas
             stars = TRUE)
```

```{r message=FALSE, warning=FALSE}
resu <- correlation(envlen4 %>%
               dplyr::rename(Length=lepro,
         SIC=meansic2,
         CHla=meanchl2,
         SSM=meantsm2))
resu
summary(resu, redundant=TRUE)

```

We can identify through a correlation matrix the data of our set,
whether it is positive or negative. See Figure \@ref(fig:coorplot).

```{r corrplot, fig.cap="Correlation Matrix", message=FALSE, warning=FALSE}
resu %>%
  summary(redundant = FALSE) %>%
  plot()
```

Here we can identify the correlation that exists between krill length
and chlorophyla (-) and krill length and SST (+)

## GLM models

Come back with `envlen2` object.

```{r}
names(envlen2)
# [1] "cellid"   "Year"     "lepro"    "meansic2" "meanchl2"
# [6] "meantsm2" "geometry"
envlen2r <- envlen2 %>% 
  rename(SIC = meansic2,
         Chl = meanchl2,
         SST = meantsm2,
         Length = lepro)

env1 <- glm(Length ~ Year,
            data = envlen2r)
env2 <- glm(Length ~ Year + 
              SIC,
            data = envlen2r)
env3 <- glm(Length ~ Year + 
              SIC +
             Chl ,
            data = envlen2r)
env4 <- glm(Length ~ Year + 
              SIC +
              Chl + 
              SST ,
            data = envlen2r)
#spatial component
env5 <- glm(Length ~ Year + 
              SIC +
              Chl +
              SST +
              cellid,
            data = envlen2r)
```

test with @Ludecke2021

Table with significance level (***).
```{r}
#summary(env4)
tab_model(env1, 
          env2,
          env3, 
          env4, p.style = "stars")

```


Table comparing performance model. 
```{r warning=F}
compare_performance(env1, 
                    env2,
                    env3,
                    env4,
                    env5,
                    env1m,
                    rank = TRUE, verbose = FALSE)
```

Regarding this result, we can select **best model** and check some assumtion like:

- Colinealidad
- Data points influyentes
- Homoscedasticidad
- Normalidad de los residuos 
- Independencia

In this case, is `env4`

```{r fig.height=3, fig.width=3,  warning=F}
check_model(env1)
```

```{r}
plot(compare_performance(env1, 
                         env2, 
                         env3,
                         env4,
                         rank = TRUE),
     name="LM to length krill data vs Env")
```

Comparative table acrosss models.

```{r}
compare_performance(env1, 
                    env2, 
                    env3,
                    env4, 
                    env5)

report(env4)
```


```{r}
AIC(env1)
AIC(env2)
AIC(env3)
AIC(env4)
deviance(env1)
deviance(env2)
deviance(env3)
deviance(env4)

library(lmtest)
lrtest0<-lrtest(env1,env2,env3,env4)
lrtest0
plot_model(env4)+
  theme_bw()
```

## GLMM models

Random effects are a way to model variability in data that comes from
factors that cannot be directly measured or controlled. In the context
of statistical models, random effects refer to variables that are
assumed to have an unknown probability distribution, and are included in
the model to explain some of the variation in the data.

For example, in a study comparing the test scores of students at
different schools, random effects refer to differences between schools
that cannot be explained by variables measured in the study, such as
students' socioeconomic status. These differences may be caused by
factors such as the quality of teaching, school culture, or geographic
location.

Random effects are often modeled by using mixed effects models, which
combine random and fixed effects in the same model. Fixed effects are
those that are assumed to be constant for all study units and are
directly measured, while random effects are those that are assumed to
vary randomly across study units and cannot be directly measured.

In short, random effects are a way of modeling variability in data that
cannot be directly explained by the variables measured in the study, and
are included in the model to improve the precision of the estimates and
reduce the potential for bias [@McCulloch2001].

In this cae we try test spatial componenten in `cellid` variable like
random effects with `lme4` package [@Bates2015].

```{r}
env1m <- lmer(Length ~ Year + 
            Chl + 
             SIC + 
             SST + (1|cellid), 
           data = envlen2r)

env2m <- lmer(Length ~ Year + 
            Chl + 
             SIC + 
             SST + (1|Year), 
           data = envlen2r)
```


```{r }
# graficar efectos aleatorios
pre <- plot_model(env1m,type = "re", 
                facet.grid=FALSE, 
                free.scale = FALSE, 
                title = NULL, 
                vline.color = "darkgrey",
                sort.est = TRUE,
                colors = "Set1",
                show.data = TRUE,
                jitter = 0.2,
                terms = "Days")+
    theme_few()

pest <- plot_model(env1m,type = "est", 
                facet.grid=FALSE, 
                free.scale = FALSE, 
                title = NULL, 
                vline.color = "darkgrey",
                sort.est = TRUE,
                colors = "Set2",
                show.data = TRUE,
                jitter = 0.2)+
  theme_few()

ggarrange(glvoid, pre, pest, ncol=3)
```


```{r warning=FALSE, message=FALSE, include=FAlSE}
ggarrange(pre, pest, ncol = 2)
```



```{r}
env_ranef <- ranef(env1m)
summary(env_ranef)
mean(env_ranef$cellid[, 1])

report(env1m)
```

# Conclusion

-   On the one hand, the models with mixed effects serve to verify the
    influence of the spatial component, in this case each cell y in
    which the data of the dependent variable (krill sizes) and the
    independent variable (environmental variables) were considered.

-   The influence of environmental variables on the sizes of the krill
    fishery is corroborated. The environmental variable with the
    greatest impact on krill sizes is Chlorophyll in negative terms. In
    other words, the more chlorophyll in the environment, the sizes
    decrease because there is greater recruitment due to the abundance
    of substrate for the krill population.

-   In a way, it is proof that the krill population structure is
    influenced not only by fishing pressure, but also by environmental
    conditions.

-   With these results, the environmental component is solidly
    incorporated into the krill stock assessment model in the Antarctic
    Peninsula, specifically in Subarea 48.1.

# References


