---
title: "Krill Length 48.1"
author: "Mauricio Mardones"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: seaice.bib
csl: apa.csl
link-citations: yes
linkcolor: blue
output:
  html_document:
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
---

```{r setup1}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

# Background

The following document intends to carry out a complementary methodological analysis to correlate environmental variables with the population dynamics of krill (*Euphausia superba*).

# Objective 

Once the correlation and effects on the population and/or fishing indicators on krill have been verified, this analysis aims to have a time series of the environmental variable to incorporate into the stock assessment process. Similar work in @Wang2021 but with a longest fishery history.



```{r message=FALSE}
library(reshape2)
library(tidyverse)
library(plyr)
library(lubridate)
library(raster)
library(sf)
library(CCAMLRGIS)
library(here)
library(easystats) # analisis estadisticos post
library(ncdf4)
library(readr)
library(data.table)
library(ggridges)
library(ggpubr)
library(knitr)
library(kableExtra)
# stat test
library(corrplot)
library(outliers)
library(visreg)
```

## Length structure krill 48 subarea

Another important piece of information for a stock evaluation refers to the biological components such as average sizes and weights across areas and years. To do this, we will explore the biological data and prepare the output to pass it to SS3.

the idea is to correlate mean lengths with SIC.

## Data exploratory analysis

The object `ohbio2` come from data exploration analysis  in data request CCAMLR data. This objetc have bio information from krill.

```{r}
#datos entregados por secretaria de CCMLAR
metadata <- load("~/DOCAS/Data/565_C1_KRI_2021-10-01/DATA_PRODUCT_565.RData")
# Data procesada por MMardones
#ohbio <- load("DataLengthKrill.RData")
#ohbio
metadata
#son lo mismo
```


```{r}
#cargo objeto
meta <- get("METADATA")
c1 <- get("C1")
ohbio <- get("OBS_HAUL_BIOLOGY")
names(ohbio)
dim(c1)
dim(ohbio)
```

Join data set  with master as `c1` set. This join is trought `obs_haul_id` variable
```{r}
ohbio2 <- left_join(c1, ohbio, by="obs_haul_id")
dim(ohbio2)
```

Firsts glance. Test how many register have by year. In this case, `length_total_cm` by season ccamlr.

same exercise in date period `date_catchperiod_start`

```{r}
ohbio3 <- ohbio2 %>%
  mutate(Year = year(date_catchperiod_start),
         Month = month(date_catchperiod_start),
         Day = day(date_catchperiod_start))
```


```{r}
names(ohbio3)
ohbio4 <- ohbio3 %>% 
dplyr::select(7, 9, 11, 12, 14, 24, 25, 29, 42, 44, 46, 47)
names(ohbio4)
```



```{r}
coutlength <-ohbio4 %>% 
  drop_na(length_total_cm) %>% 
  dplyr::group_by(Year, Month, asd_code) %>%
  dplyr::summarise(avg=mean(length_total_cm))
#kableExtra::kable(coutlength, format = "html")
```

Some features of data set length structures
```{r}
table(ohbio4$gear_type) %>%
  kbl() %>%
  kable_styling()

table(ohbio2$trawl_technique)%>%
  kbl() %>%
  kable_styling()
```

## Plots

Filter data regarding to previous glances. Follow with a quick glimse to all 48 subarea length composition from monitoring fisheries.

```{r  fig.align="center", message=F, warning=F}
jz <- ggplot(ohbio4 %>% 
               filter(Year>2000),
             aes(x=length_total_cm, 
                 y = as.factor(Year), 
                 fill=asd_code))+
  #geom_joy(alpha=0.9) +
  geom_density_ridges(stat = "binline", bins = 50, 
                      scale = 1.8, 
                      draw_baseline = FALSE,
                      alpha=0.9)+
  facet_wrap(.~asd_code, ncol=3) +   
  geom_vline(xintercept = 3.6, color = "red")+
  scale_x_continuous(breaks = seq(from = 1, to = 10, 
                                  by = 1))+
  scale_y_discrete(breaks = seq(from = 2000, 
                                to = 2020, by = 1))+
  scale_fill_viridis_d(name="SubArea",
                       option="F")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))+
  xlim(0,10)+
  xlab("Longitud (cm.)")+
  ylab("")
jz
```

```{r  fig.align="center", message=F, warning=F}

trwal <- ggplot(ohbio4 %>% 
               filter(Year>2000),
             aes(x=length_total_cm, 
                 y = as.factor(Year),
                 fill= trawl_technique))+
  #geom_joy(alpha=0.9) +
  geom_density_ridges(stat = "binline", bins = 50, 
                      scale = 1.8, 
                      draw_baseline = FALSE,
                      alpha=0.5)+
  facet_wrap(.~asd_code, ncol=3) +   
  geom_vline(xintercept = 3.6, color = "red")+
  scale_x_continuous(breaks = seq(from = 1, to = 10, 
                                  by = 1))+
  scale_y_discrete(breaks = seq(from = 2000, 
                                to = 2020, by = 1))+
  scale_fill_viridis_d(name="Trawl Technique",
                       option="F")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))+
  xlim(0,10)+
  xlab("Longitud (cm.)")+
  ylab("")
trwal
```
Same plot by Sub Area and month trought year. With this kind of plot, we can see the recruit power interannually.


```{r  fig.align="center", message=F, warning=F}
ymasd <- ggplot(ohbio4 %>% 
               filter(Year>2000),
             aes(x=length_total_cm, 
                 y = as.factor(Month),
                 fill=asd_code))+
  #geom_joy(alpha=0.9) +
  geom_density_ridges(stat = "binline", bins = 50, 
                      scale = 1.5, 
                      draw_baseline = FALSE,
                      alpha=0.5)+
  facet_wrap(.~season_ccamlr, ncol=5) +   
  geom_vline(xintercept = 3.6, color = "red")+
  scale_x_continuous(breaks = seq(from = 1, to = 10, 
                                  by = 1))+
  scale_y_discrete(breaks = seq(from = 1, 
                                to = 12, by = 1))+
  scale_fill_viridis_d(name="SubArea",
                       option="F")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))+
  xlim(0,10)+
  xlab("Longitud (cm.)")+
  ylab("")
ymasd
```

```{r  fig.align="center", message=F, warning=F}
yma481 <- ggplot(ohbio4 %>% 
               filter(Year>2000,
                      asd_code==481),
             aes(x=length_total_cm, 
                 y = as.factor(Month),
                 fill=asd_code))+
  #geom_joy(alpha=0.9) +
  geom_density_ridges(stat = "binline", bins = 50, 
                      scale = 1.5, 
                      draw_baseline = FALSE,
                      alpha=0.5)+
  facet_wrap(.~Year, ncol=5) +   
  geom_vline(xintercept = 3.6, color = "red")+
  scale_x_continuous(breaks = seq(from = 1, to = 10, 
                                  by = 1))+
  scale_y_discrete(breaks = seq(from = 1, 
                                to = 12, by = 1))+
  scale_fill_viridis_d(name="SubArea",
                       option="F")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, 
                                   hjust = 1))+
  xlim(0,10)+
  xlab("Longitud (cm.)")+
  ylab("")
yma481
```

## Mean Length
Recruit estimate (% Low Recruit length) [@Perry2020]


Un diaggrama de caja para el porcentage de individuos bajo talla (3.6 mmm.)
```{r}
box <- ggplot(ohbio4) +
  geom_boxplot(aes(length_total_cm, group=season_ccamlr), 
               alpha=0.3,
               fill=3)+
  geom_vline(xintercept = 3.6, color = "red")+
  xlim(0, 7)+
  coord_flip()+
  facet_wrap(~asd_code, ncol=1)+
  theme_bw()+
  labs(x="% Below Recruit length (3.6 mm.)",
       x="")
hi <- ggplot(ohbio4)+
  geom_histogram(aes(y=length_total_cm), 
                 fill=3,
                 alpha=0.3,
                 color="black")+
  coord_flip()+
  theme_bw()+
  ylim(0,10)

box | hi
  
```

Ahora saco tallas medias por distintas variables


```{r, fig.align="center"}
pmea <- ggplot(coutlength %>% 
                 filter(Year>2000), 
               aes(Year,avg))+
    geom_point(shape=21, fill=coutlength$asd_code, 
               show.legend = T) +
    stat_smooth(method= "glm", colour='#253494')+
    scale_size(range = c(-4,8)) +
    theme_bw()+ 
    facet_wrap(.~asd_code)+
    scale_x_continuous(breaks = seq(from = 2000, to = 2020, by = 2))+
    #scale_y_discrete(breaks = seq(from = 1, to = 13, by = 1))+
    theme(axis.text.x = element_text(angle = 90, hjust = 2))+
    guides(fill = guide_legend(reverse=F))+
    scale_fill_viridis_c(option="E")+
    ylim(3,6)+
    ylab("") +
    xlab("") +
    ggtitle("Lenght Mean Krill fishery")
pmea
```
by month and year only 48.1


```{r, fig.align="center"}
pmea481 <- ggplot(coutlength %>% 
                 filter(Year>2000,
                        asd_code==481), 
               aes(Month, avg))+
    geom_point(shape=21,  
               show.legend = T) +
    stat_smooth(method= "glm", colour='#253494')+
    scale_x_continuous(breaks = seq(from = 1, to = 12, by = 2))+
    theme_bw()+ 
    facet_wrap(.~Year) +
    ggtitle("Lenght Mean Krill fishery 48.1")+
    ylim(0, 6)
pmea481
```


## GLM just length

```{r}
model1 <- glm(avg ~ Year, data = coutlength)
r2(model1)
check_model(model1)

model2 <- glm(avg ~ Year + 
                Month, data = coutlength)
r2(model2)
check_model(model2)

model3 <- glm(avg ~ Year + 
                Month + 
                asd_code, data = coutlength)
r2(model3)
check_model(model3)
```

Compara modelos

```{r}
plot(compare_performance(model1, 
                         model2, 
                         model3,
                         rank = TRUE))
```

(Probar datos con RaadTools library [@Raadtools2022])

## Grid to data

In this step we filter the empty or null data (`NA`) and we will nest the mean lengths in the same formats as the environmental data.


```{r agrupa}

# Cargo linea de costa
coast <- load_Coastline()
coast1<- st_as_sf(coast) 
coast2 = st_transform(coast1, "+proj=latlong +ellps=WGS84")
# con SSMU
ssmu <- load_SSMUs()
ssmu481 <- subset(ssmu[c(2,3,4,5,6,7,17),])
ssmu481a <- st_as_sf(ssmu481) 
ssmu481aa = st_transform(ssmu481a, "+proj=latlong +ellps=WGS84")

# con Statistical Areas con foco en 48.1
suba <- load_ASDs()
suba1 <- subset(suba[(3),])
suba1a<- st_as_sf(suba1) 
suba1aa = st_transform(suba1a, "+proj=latlong +ellps=WGS84")


# Uso las agrupaciones de DMP1
dmp1 <- st_read("~/DOCAS/Mapas/Antarctic_SHPfiles/D1MPA-model-2019.shp",
                quiet=T)
dmp1=st_transform(dmp1, 6932)

# Uso las agrupaciones de Strata
strata <- st_read("~/DOCAS/Mapas/Antarctic_SHPfiles/Strata.shp",
                quiet=T)
strata=st_transform(strata, "+proj=latlong +ellps=WGS84")
```

```{r maptest}
# y testeo el mapa
ssmap <- ggplot(ssmu481aa)+
  geom_sf(data = strata, aes(fill=strata$ID, 
                           alpha=0.3))+
  # geom_sf(data = ssmu481aa, aes(fill=ssmu481aa$GAR_Short_Label, 
  #                         alpha=0.3))+
  geom_sf(data = coast2, colour="black", fill=NA)+
  #geom_sf(data = gridcrop1, colour="black", fill=NA)+
  #geom_sf(data= suba1aa, fill=NA)+
  # geom_sf(aes(fill=ssmu481aa$GAR_Short_Label,
  #              alpha=0.3))+
  scale_fill_viridis_d(option = "F",
                       name="Strata")+
  #geom_sf_label(aes(label = strata$ID))+
  # labs(fill = "SSMU")+
  ylim(230000, 2220000)+
  xlim(-3095349 , -1858911)+
  # coord_sf(crs = 32610)+ #sistema de prpyecccion para campos completos
  coord_sf(crs = 6932)+
  scale_alpha(guide="none")+
  theme_bw()
ssmap
```
## Grouping Length data

## SSMU 48.1 CCAMLR

Identificar tendencias temporales por cada SSSMU para tener un mejor detalle del comportamiento de la variable a traves del tiempo en el espacio definido.

De forma experimental hago una grilla sobre la 48.1.

```{r ssmu1}
names(ohbio4)
ohbio5 <- ohbio4 %>% 
  filter(asd_code==481) %>% 
  dplyr::select(6, 7, 8, 9, 10, 11, 12) 
ohbio6 <- st_as_sf(ohbio5 %>% 
                     drop_na(latitude_set_end), 
                   coords = c("longitude_set_end", "latitude_set_end"),  
                  crs = "+proj=latlong +ellps=WGS84")
```

hago la grilla
```{r ssmu1}

Grid<- suba1aa  %>% #pm481 es el plot base original linea 481
  sf::st_make_grid(cellsize = c(1,0.5)) %>% # para que quede cuadrada
  sf::st_cast("MULTIPOLYGON") %>%
  sf::st_sf()  %>%  # objeto en spatial feature
  dplyr::mutate(cellid = row_number()) 

# Clean the input data by removing duplicate vertices and making the object topologically valid
grid3 <- st_make_valid(Grid)

# Corto la grilla dentro de las SSMU
gridcrop1 <- crop_shape(grid3, suba1aa, polygon = TRUE)

# the first object drives the output geometry
grilen <- grid3 %>%
  st_join(ohbio6) %>% 
  group_by(cellid)

names(grilen)
dim(grilen)
class(grilen)
head(grilen)

```

Test lentgh mean map by Year

```{r ssmugri}
grilen3 <- grilen %>%
  drop_na(length_total_cm) %>% 
  dplyr::group_by(cellid,Year) %>% 
  dplyr::summarise(lepro=mean(length_total_cm))

 
gl <- ggplot() +
   geom_sf(data=grilen3,
           aes(fill = lepro), 
           alpha=0.7, color=NA) +
  scale_fill_viridis_b(option="G",
                       name="Krill Length Fishery (cm.)",
                       breaks = seq(0,8,1),
                       direction=-1)+
  geom_sf(data = suba1aa, fill=NA, col=2)+
  geom_sf(data = coast2, colour="black", fill="grey")+
  facet_wrap(~Year, ncol=6)+
  theme(panel.background = element_rect(fill = 'aliceblue'),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
  # ylim(-70, -60)+
  # xlim(-70, -50)
  # another kind projection
  ylim(230000, 2220000)+
  xlim(-3095349 , -1858911)+
  #coord_sf(crs = 32610)+ #sistema de prpyecccion para campos completos
  coord_sf(crs = 6932)
gl
```

Load environmental data to merge with length.

```{r eval=FALSE}
datae <- load("~/DOCAS/Data/KrillLows/DataEnvKrill.RData")
```


```{r}
ls()
get("dataenvf")
```

Join length and environental data by `cellid` and `Year` with a reduce DB. 
```{r}
class(dataenvf)

envgru <- dataenvf %>%
  dplyr::select(1, 2, 3, 6, 9) %>% 
  drop_na(ANO.x) %>% 
  dplyr::rename(Year=ANO.x) %>% 
  dplyr::group_by(cellid,Year) %>% 
  dplyr::summarise(meansic2=mean(meansic),
                   meanchl2=mean(meanchl),
                   meantsm2=mean(meantsm))
dim(envgru)
summary(envgru)

```
Join merge object in a new data base.

```{r}
names(grilen3)
names(envgru)
envlen <- merge(grilen3, envgru, 
                 by = c("cellid", "Year"))
```

## GLM Length-Environmental

Debo trabajar ciertos componentes de los datos para poder hacer el analisis bajo supuestos estructurales.

- Comprobar distribución de Longitudes medias de krill
- Cambiar a factor `Year` y `cellid`
- Filtrar datos `NA`
- Filtrar outliers en datos ambientales


### Test de comportamiento de la variable de longitudes `lenpro` mediante 2 test; 1) Shapiro-Wilk `shapiro.test()` y Kolmogorov-Smirnov `ks.test()`

```{r}
shapiro.test(envlen$lepro)
ks.test(envlen$lepro, "pnorm", mean(envlen$lepro), sd(envlen$lepro))
tal<-ggplot(envlen, aes(x=lepro)) + 
  geom_histogram(color="red", fill="white")+
  theme_minimal()+
  labs(x="Length Mean Krill (cm.)",
       y="Frecuency")
tal
```
### Cambiar a factor `Year` y `cellid`

```{r}
envlen$Year<-as.factor(envlen$Year)
envlen$cellid<-as.factor(envlen$cellid)
glimpse(envlen)
```
### Filtrar datos `NA`

```{r}
# Are there missing values?
colSums(is.na(envlen))
```
### Filter outliers environmental and biological data

Test trought plot all variables in `envlen` object
```{r}
boxsic <- ggplot(envlen)+
  geom_boxplot(aes(Year, meansic2),
               outlier.colour = "red")+
  theme_minimal()
boxtsm <- ggplot(envlen)+
  geom_boxplot(aes(Year, meantsm2),
               outlier.colour = "red")+
  theme_minimal()
boxchl <- ggplot(envlen)+
  geom_boxplot(aes(Year, meanchl2),
               outlier.colour = "red")+
  theme_minimal()
ggarrange(boxsic, boxtsm, boxchl, nrow = 3)

```
got off sic outliers data and again

```{r}
envlen2 <- envlen %>% 
  filter(meansic2<100)

#plot 2
         
boxsic2 <- ggplot(envlen2)+
  geom_boxplot(aes(Year, meansic2),
               outlier.colour = "red")+
  theme_minimal()
boxtsm2 <- ggplot(envlen2)+
  geom_boxplot(aes(Year, meantsm2),
               outlier.colour = "red")+
  theme_minimal()
boxchl2 <- ggplot(envlen2)+
  geom_boxplot(aes(Year, meanchl2),
               outlier.colour = "red")+
  theme_minimal()

p<-ggplot(envlen2, aes(x=meansic2)) + 
  geom_histogram(color="black", fill="white")+
  theme_minimal()

q<-ggplot(envlen2, aes(x=meantsm2)) + 
  geom_histogram(color="black", fill="white")+
  theme_minimal()

t<-ggplot(envlen2, aes(x=meanchl2)) + 
  geom_histogram(color="black", fill="white")+
  theme_minimal()

ggarrange(boxsic2, boxtsm2, boxchl2, p, q, t, nrow = 2, ncol=3)
```
A su vez un plo t de hist 

# Correlation Analysis 

## Simple correlation identification

```{r}
class(envlen3)
envlen3 <- as.data.frame(envlen2)
report(envlen4)
envlen4 <- envlen3 %>% 
  dplyr::select(3, 4, 5, 6) %>% 
  
library(MASS)


plot(envlen2$meantsm2, envlen2$lepro, pch = 19, col = "lightblue")
abline(lm(envlen2$lepro ~ envlen2$meantsm2), col = "red", lwd = 3)
# Correlación de Pearson
text(paste("Correlación:", round(cor(envlen2$meantsm2, envlen2$lepro), 2)), x = 272, y = 5)

library(PerformanceAnalytics)

chart.Correlation(envlen4, histogram = TRUE, method = "pearson")

library(psych)

pairs.panels(envlen4,
             smooth = TRUE,      # Si TRUE, dibuja ajuste suavizados de tipo loess
             scale = FALSE,      # Si TRUE, escala la fuente al grado de correlación
             density = TRUE,     # Si TRUE, añade histogramas y curvas de densidad
             ellipses = TRUE,    # Si TRUE, dibuja elipses
             method = "spearman", # Método de correlación (también "spearman" o "kendall")
             pch = 21,           # Símbolo pch
             lm = FALSE,         # Si TRUE, dibuja un ajuste lineal en lugar de un ajuste LOESS
             cor = TRUE,         # Si TRUE, agrega correlaciones
             jiggle = FALSE,     # Si TRUE, se añade ruido a los datos
             factor = 2,         # Nivel de ruido añadido a los datos
             hist.col = 4,       # Color de los histogramas
             stars = TRUE,       # Si TRUE, agrega el nivel de significación con estrellas
             ci = TRUE)          # Si TRUE, añade intervalos de confianza a los ajustes

report(rep)
library(psych)

corPlot(envlen2, cex = 1.2, main = "Matriz de correlación")
ma <- cor(envlen4)
reg3 <- glm(lepro ~ meantsm2 + meanchl2 + meansic2, 
            family = gaussian,
            data = envlen)
plot(reg3)
summary(reg3)

visreg(reg3, "meanchl2", partial = F)
qqPlot(reg3)
```

```{r}
boxp <- ggplot(envlen %>% 
                 filter(meansic2<100))+
  geom_boxplot(aes(Year, meansic2))+
  theme_minimal()
boxp

envlenhis <- envlen %>% 
  filter(meansic2<100) 

hist(log(envlenhis$meansic2))

# Identificar y sacar los outliers utilizando el método de Tukey
envsic_clean <- rm.outlier(envlen$meansic2, fill = TRUE, median = FALSE)
```



```{r}
env1 <- glm(lepro ~ Year,
            data = envlen)
r2(env1)
summary(env1)
check_model(env1)
check_heteroscedasticity(env1)
model_performance(env1)
drop1(env1)
anova(env1, test="Chi")
summary(env1)
```


```{r}
library(lme4)
m3 <- lmer(lepro ~ Year + (1 + Year | meantsm2), data = envlen)
model_performance(m3)

model2 <- glm(avg ~ Year + 
                Month, data = coutlength)
r2(model2)
check_model(model2)

model3 <- glm(avg ~ Year + 
                Month + 
                asd_code, data = coutlength)
r2(model3)
check_model(model3)
```


# References

Sea ice data is from Giovanni NASA browser 
(https://giovanni.gsfc.nasa.gov/)
ref: Global Modeling and Assimilation Office (GMAO) (2015), 

MERRA-2 tavgM_2d_flx_Nx: 2d,Monthly mean,Time-Averaged,Single-Level,
Assimilation,Surface Flux Diagnostics V5.12.4, 
Greenbelt, MD, USA, 
Goddard Earth Sciences Data and Information Services Center (GES DISC), 
Accessed: [06 July 2022], 10.5067/0JRLVL8YV2Y4